---
title: "AstroStats Final Project"
author: "Mira Flynn"
date: "4/6/2022"
output: 
  github_document:
    toc: TRUE
---

## Setup and Data Cleaning

```{r setup}
# Dependencies
library(tidyverse)
library(magrittr)
library(sf)

# Create the final PDF with code snippets shown
knitr::opts_chunk$set(echo = TRUE)

sf::sf_use_s2(FALSE) # this fixes the geometry of the shapefile for some reason


# This is a premade graph theme I originally got from ZDR and modified for myself
theme_common <- function() {
  theme_minimal() %+replace%
  theme(
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    axis.title.x = element_text(margin = margin(4, 4, 4, 4), size = 16),
    axis.title.y = element_text(margin = margin(4, 4, 4, 4), size = 16, angle = 90),
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 12),
    strip.text.x = element_text(size = 12),
    strip.text.y = element_text(size = 12),
    panel.grid.major = element_line(color = "grey90"),
    panel.grid.minor = element_line(color = "grey90"),
    aspect.ratio = 10/16,
    plot.margin = unit(c(t = +0.5, b = +0, r = +0, l = +0), "cm"),
    plot.title = element_text(size = 18),
    # plot.title.position = "plot",
    plot.subtitle = element_text(size = 16),
    plot.caption = element_text(size = 12)
  )
}

## Helper function to compute uncertainty bounds
# More taking code from ZDR!
add_uncertainties <- function(data, model, prefix = "pred", ...) {
  df_fit <-
    stats::predict(model, data, ...) %>%
    as_tibble() %>%
    rename_with(~ str_c(prefix, "_", .))

  bind_cols(data, df_fit)
}

rad2deg <- function(rad) {(rad * 180) / (pi)}
deg2rad <- function(deg) {(deg * pi) / (180)}
# https://stackoverflow.com/questions/32370485/convert-radians-to-degree-degree-to-radians

# r = Radius of Moon in KM
r <- 1737.4

# How much to round the latitudes. Round to spacing of 5
round_digits = 5

```

This is all the data cleaning that I did on the Robbins dataset. It is mainly working with the shapefile to categorize each point by biome.
```{r data-cleaning}
map = read_sf("data/LROC_GLOBAL_MARE_180.SHP") %>% # Load the maria shapefile
  st_make_valid() # I don't know what this does exactly but it fixes the shapefile
st_crs(map$geometry) <- 4326 # Again, I don't understand

df_craters <- read.csv("data/AstroStats_Robbins_Moon.csv") %>% # Load the dataset
  mutate( # Change longitude from 0 to 360  -> -180 to 180
      LON_ELLI_IMG = ifelse(LON_ELLI_IMG > 180, LON_ELLI_IMG-360, LON_ELLI_IMG)
  ) %>%
  filter(!is.na(LAT_ELLI_IMG)) %>% # Get rid of the 8 craters with NA ellipses
  st_as_sf( # Convert the craters dataset to a shapefile object
    coords = c("LON_ELLI_IMG", "LAT_ELLI_IMG"), 
    crs = st_crs(map),
    remove = FALSE
  ) %>% 
  mutate(
    intersection = geometry %>% st_intersects(map, sparse = TRUE),
    # Intersection column is where each point intersects the map, if any
    MARIA = (intersection %>% as.character()) != "integer(0)"
    # This is a bit chaotic. The intersection column is a funky object type, so
    # to compare it, I cast it to a character type. If the intersection is 0,
    # then the point is not in one of the maria regions. 
  ) %>%
  as.data.frame() %>% # Need to cast back to dataframe to get rid of geometry
  select( # Get rid of the intersection and geometry columns for clarity.
    -c(intersection, geometry)
  )

df_craters %>% write.csv(
    "data/AstroStats_Robbins_Moon_Maria.csv", 
    row.names = FALSE
  ) # Write modified data to CSV
```


```{r data-overview}
df_craters %>% head()
df_craters %>% summary()
```
## Initial looks at several variables

Note: These plots were intended for data exploration, and therefore are not presentation quality. 

```{r angle-variation}
df_craters %>%
  ggplot(aes(x = LAT_ELLI_IMG, y = DIAM_ELLI_ANGLE_IMG)) +
  geom_point(alpha = 0.1) +
  geom_smooth()
```

This graph explores the angle for the most eccentric craters. Craters with low eccentricity are nearly circular, so the direction they point may be more random.

```{r eccentric-craters-angle-variation}
df_craters %>%
  mutate(
    LAT_ELLI_IMG = abs(LAT_ELLI_IMG),
    DIAM_ELLI_ANGLE_IMG = DIAM_ELLI_ANGLE_IMG %% 180
  ) %>%
  filter(DIAM_ELLI_ECCEN_IMG > 0.8) %>%
  filter(LAT_ELLI_IMG < 60) %>%
  ggplot(aes(x = LAT_ELLI_IMG, y = DIAM_ELLI_ANGLE_IMG)) +
  geom_point(alpha = 0.1) +
  geom_smooth()
```

This graph seems to show some eccentricity variation across different latitudes, but there are noticeable lines where the Robbins paper switched projections. Therefore, I decided not to explore this pattern further.

```{r eccentricity-variation}
df_craters %>%
  ggplot(aes(x = LAT_ELLI_IMG, y = DIAM_ELLI_ECCEN_IMG)) +
  geom_point(alpha = 0.1) +
  geom_smooth()
```

```{r major-diameter-variation}
df_craters %>%
  ggplot(aes(x = LAT_ELLI_IMG, y = DIAM_ELLI_MAJOR_IMG)) +
  geom_point(alpha = 0.1) +
  geom_smooth() +
  scale_y_log10()
```


```{r minor-diameter-variation}
df_craters %>%
  ggplot(aes(x = LAT_ELLI_IMG, y = DIAM_ELLI_MINOR_IMG)) +
  geom_point(alpha = 0.1) +
  geom_smooth() +
  scale_y_log10()
```

Plotting craters by latitude and longitude creates essentially a map of the moon.

```{r map}
df_craters %>%
  ggplot(aes(x = LON_ELLI_IMG, y = LAT_ELLI_IMG)) +
  geom_point(alpha = 0.1) +
  coord_fixed()

# WOAH THIS IS SO COOL LOOK AT THIS
# https://www.physics.unlv.edu/~jeffery/astro/moon/map/moon_map_mercator.jpg
```


This histogram was my first look at density, and is affected by the circumference of the moon being different at each latitude

```{r density-histogram}
df_craters %>%
  ggplot(aes(LAT_ELLI_IMG)) +
  geom_histogram(bins = 36) +
  scale_x_continuous(breaks = c(-90,-60,-30,0,30,60,90))

# NOTE: This isn't accounting for diameter changes over latitudes
```

## Shapefile Work

This is the equally spaced points data, which I used to calculate area of each biome at each latitude

```{r equally-spaced-points}
# Create empty lists of lats and lons
lats = c()
lons = c()

# Add every combo of lats and lons, spaced at 2 degrees, to the list
for (lat in seq(-90,90,1)){
  for (lon in seq(-180,180,1)){
    lats = c(lats, lat)
    lons = c(lons, lon)
  }
}


pnts <- data.frame(lat = lats, lon = lons) %>% # Make a dataframe of the evenly
  # spaced points
  st_as_sf( # Then do the same things as we did with the craters dataset.
    coords = c("lon", "lat"), 
    crs = st_crs(map),
    remove = FALSE
  ) %>% 
  mutate(
    intersection = geometry %>% st_intersects(map) %>% as.integer(),
    # For some reason, this dataframe works a bit differently. I don't
    # understand why, and I'm not gonna spend a bunch of time figuring out why
    maria = !is.na(intersection)
  )

pnts %>% summary()
pnts %>%
  ggplot() +
  geom_point(aes(x = lon, y = lat, color = maria), size = 0.2) +
  coord_fixed()

pnts %>% write.csv(
    "data/AstroStats_Degree_Spaced_Maria_Or_Highlands.csv", 
    row.names = FALSE
  ) # Write modified data to CSV
```

This calculates the portion of the area at each latitude that is in the maria.

```{r maria-portions}
maria_portions <- pnts %>%
  mutate(
    LAT_ROUND = round(lat/round_digits)*round_digits # Round to nearest latitude bin
  ) %>%
  filter(
    abs(LAT_ROUND) < 90 # 0 circumference at 90 degrees
  ) %>%
  group_by(LAT_ROUND) %>%
  summarize(
    PORTION_MARIA = mean(maria) # Get the mean of the true/false maria column
  ) %>%
  as.data.frame() %>%
  select(c(-geometry))
maria_portions %>% head()
```



```{r biome-area}
df_craters_hist <- df_craters %>%
  mutate(
    LAT_ROUND = round(LAT_ELLI_IMG/round_digits)*round_digits
  ) %>%
  filter(
    abs(LAT_ROUND) < 90 # Get rid of 90s because they round to 0 area
  ) %>%
  group_by(LAT_ROUND, MARIA) %>%
  summarize(
    COUNT_LAT_ROUND = n() # Count number of craters in each biome and latitude
  ) %>%
  mutate(
    CIRCUMFERENCE_LAT = 2*pi*r*cos(deg2rad(LAT_ROUND)),
    NORM_LAT_ROUND = COUNT_LAT_ROUND/CIRCUMFERENCE_LAT # Normalize by area
  ) %>%
  left_join(maria_portions, by = "LAT_ROUND") %>%
  mutate(
    CIRCUMFERENCE_BIOME = ifelse(
      MARIA, 
      CIRCUMFERENCE_LAT * PORTION_MARIA, 
      CIRCUMFERENCE_LAT * (1-PORTION_MARIA)),
    BIOME = ifelse(MARIA, "MARIA", "HIGHLANDS") # Get the final circumference 
    # for each biome and latitude
  )
df_craters_hist %>% head()
```


## Processed Density Graphs

This histogram is recreating the histogram I did way above, to check that my work didn't mess anything up

```{r basic-histogram}
df_craters_hist %>%
  ggplot(aes(x = LAT_ROUND, y = COUNT_LAT_ROUND)) +
  geom_col(
    aes(
      # fill = LAT_ROUND %>% as.character() %>% reorder(LAT_ROUND)
    ), 
    width = round_digits, 
    color = "black", 
    size = 1,
    show.legend = FALSE
  ) +
  scale_x_continuous(
    breaks = seq(-90,90,15),
    minor_breaks = seq(-90,90,5)
  ) +
  # scale_y_log10() +
  theme_common() +
  theme(
    panel.grid.major = element_line(color = "darkgrey"),
    panel.grid.minor = element_line(color = "grey90"),
  ) +
  labs(
    x = "Latitude (degrees)",
    y = "Craters"
  )
  
```

This is the normalized density over the whole lunar surface

```{r normalized-density-histogram}
df_craters_hist %>%
  group_by(LAT_ROUND) %>%
  summarize(
    NORM_LAT_ROUND = sum(COUNT_LAT_ROUND)/mean(CIRCUMFERENCE_LAT)
  ) %>%
  ggplot(aes(x = LAT_ROUND, y = NORM_LAT_ROUND)) +
  geom_col(
    aes(
      # fill = LAT_ROUND %>% as.character() %>% reorder(LAT_ROUND)
    ), 
    width = round_digits, 
    color = "black", 
    size = 1,
    show.legend = FALSE
  ) +
  scale_x_continuous(
    breaks = seq(-90,90,15),
    minor_breaks = seq(-90,90,15)
  ) +
  # scale_y_log10() +
  theme_common() +
  theme(
    panel.grid.major = element_line(color = "grey90"),
    # panel.grid.minor = element_line(color = "grey90"),
  ) +
  labs(
    x = "Latitude (degrees)",
    y = "Crater Density \n (craters/km)",
    # caption = "\n Lunar crater density distribution at different latitudes \n"
  )
```

Normalized density for each biome, with bounds of the shapefile indicated

```{r biome-normalized-density-histogram}

df_craters_hist %>%
  mutate(
    NORM_LAT_ROUND_BIOME = COUNT_LAT_ROUND/CIRCUMFERENCE_BIOME
  ) %>%
  
  ggplot(aes(x = LAT_ROUND, y = NORM_LAT_ROUND_BIOME)) +
  geom_col(
    aes(
      # fill = LAT_ROUND %>% as.character() %>% reorder(LAT_ROUND)
    ), 
    width = round_digits, 
    color = "black", 
    size = 1,
    show.legend = FALSE
  ) +
  # geom_vline(aes(xintercept = 64.7, color = "Bounds of Maria Shapefile")) +
  # geom_vline(aes(xintercept = -66.5, color = "Bounds of Maria Shapefile")) +
  geom_vline(xintercept = 64.7) +
  geom_vline(xintercept = -66.5) +
  facet_wrap(vars(BIOME)) +
  scale_x_continuous(
    breaks = seq(-90,90,30),
    minor_breaks = seq(-90,90,15)
  ) +
  # scale_y_log10() +
  theme_common() +
  theme(
    # panel.grid.major = element_line(color = "darkgrey"),
    panel.grid.major = element_line(color = "grey90"),
    # panel.grid.minor = element_line(color = "grey90"),
  ) +
  labs(
    x = "Latitude (degrees)",
    y = "Crater Density By Biome \n (Craters/KM)"
  )
```
Mean and SD of the density over the whole lunar surface.
```{r mean-sd}
df_craters_hist %>%
  group_by(LAT_ROUND) %>%
  summarize(
    COUNT_LAT_ROUND = sum(COUNT_LAT_ROUND),
    CIRCUMFERENCE_LAT = mean(CIRCUMFERENCE_LAT),
    NORM_LAT_ROUND = COUNT_LAT_ROUND/CIRCUMFERENCE_LAT
  ) %>%
  summarize(
    mean = mean(NORM_LAT_ROUND),
    sd = sd(NORM_LAT_ROUND)
  )
```


Mean and SD of the density for each biome

```{r mean-sd-biomes}
df_craters_hist %>%
  mutate(
    NORM_LAT_ROUND_BIOME = COUNT_LAT_ROUND/CIRCUMFERENCE_BIOME
  ) %>%
  group_by(BIOME) %>%
  summarize(
    mean = mean(NORM_LAT_ROUND_BIOME),
    sd = sd(NORM_LAT_ROUND_BIOME)
  )
```


## More stuff I'm leaving in

Here is the crater map, but categorized by biome
```{r categorized-craters}
df_craters %>%
  ggplot() +
  geom_point(
    aes(x = LON_ELLI_IMG, y = LAT_ELLI_IMG, color = MARIA), 
    size = 1,
    alpha = 0.2) +
  coord_fixed()


```

The original shapefile

```{r shapefile}
# Showing the original shapefile
map %>%
  ggplot() +
  geom_sf(
    fill = "blue",
    color = "blue"
    )
```


Portion of maria across different latitudes
```{r maria-portion}
df_craters_hist %>%
  filter(MARIA == FALSE) %>%
  ggplot(aes(x = LAT_ROUND, y = 1-PORTION_MARIA)) +
  geom_col(
    aes(
      # fill = LAT_ROUND %>% as.character() %>% reorder(LAT_ROUND)
    ), 
    width = round_digits, 
    color = "black", 
    size = 1,
    show.legend = FALSE
  ) +
  scale_x_continuous(
    breaks = seq(-90,90,15),
    minor_breaks = seq(-90,90,5)
  ) +
  # scale_y_log10() +
  theme_common() +
  theme(
    panel.grid.major = element_line(color = "darkgrey"),
    panel.grid.minor = element_line(color = "grey90"),
  ) +
  labs(
    x = "Latitude (degrees)",
    y = "Maria Portion per Latitude"
  )
```

The density vs. the portion of maria, with points for each latitude. This shows a relatively linear trend with more maria meaning lower density

```{r density-vs-maria}
df_craters_hist %>%
  mutate(LAT_ROUND = abs(LAT_ROUND)) %>%
  group_by(LAT_ROUND) %>%
  summarize(
    COUNT_LAT_ROUND = sum(COUNT_LAT_ROUND),
    CIRCUMFERENCE_LAT = mean(CIRCUMFERENCE_LAT),
    PORTION_MARIA = mean(PORTION_MARIA)
  ) %>%
  mutate(
    COUNT_LAT_ROUND = ifelse(LAT_ROUND == 0, COUNT_LAT_ROUND * 2, COUNT_LAT_ROUND),
    NORM_LAT_ROUND = COUNT_LAT_ROUND/CIRCUMFERENCE_LAT
  ) %>%
  ggplot(aes(x = LAT_ROUND, y = NORM_LAT_ROUND)) +
  geom_point()
```




oh no there were so many links I used

https://wms.lroc.asu.edu/lroc/view_rdr/SHAPEFILE_LROC_GLOBAL_MARE

https://gis.stackexchange.com/questions/133625/checking-if-points-fall-within-polygon-shapefile

https://r-spatial.org/r/2017/03/19/invalid.html

https://gis.stackexchange.com/questions/413584/fix-features-with-invalid-spherical-geometry-polygon-shape-file-for-s2-geometr

https://stackoverflow.com/questions/54734771/sf-write-lat-long-from-geometry-into-separate-column-and-keep-id-column

https://stackoverflow.com/questions/54734771/sf-write-lat-long-from-geometry-into-separate-column-and-keep-id-column

